\Entry{Explainer}{\string\Explainer}{@BExplainer}
\DeclareTerm*{\string\Explainer}{@BExplainer}

The Tectonopedia command \b{\string\Explainer} is used to declare a new output
representing an “explainer” document.

\section*{Usage}

\begin{texdisp}
\Explainer{SLUG}{TEX-TITLE}{PLAIN-TITLE}
\end{texdisp}

The \tex`SLUG` is the unique URL path to this explainer. The explainer will be
located at the path \tex`/explain/{SLUG}`. Every explainer's slug must be unique.

The \tex`TEX-TITLE` is the title to be used for the explainer, including
\TeX/\LaTeX\ markup. This title will appear at the top of the resulting output
page.

The \tex`PLAIN-TITLE` is the explainer title in plain Unicode, without any
\TeX/\LaTeX\ markup. It is parsed in verbatim mode, so that control sequences
and other special characters do not have their usual \TeX\ meanings. It will be
used in places such as metadata records where \TeX/\LaTeX\ typesetting cannot be
used.

\section*{Example}

\begin{texdisp}
\explainer{why-tex}{Why \TeX?}{Why TeX?}

Technology has come a long way since the 1970's ...
\end{texdisp}


\Entry{explain}{\string\explain}{@Bexplain}
\DeclareTerm{@Bexplain}

The Tectonopedia \b{\string\explain} command creates a link to the specified
explainer page.

\section*{Usage}

\begin{texdisp}
\explain{SLUG}
\end{texdisp}

The \tex`SLUG` is the slug of the explainer to reference.

In the output, this command expands to a hyperlink to the specified explainer whose
text is the entry title.

\section*{Example}

\begin{texdisp}
For more justification, see the \explain{why-tex} page.
\end{texdisp}
