\Entry{href}{\string\href}{\href}
\DeclareTerm*{\string\href}{@Bhref}

The Tectonopedia command \b{\string\href} creates a hyperlink. It is intended to
be used for outgoing links that take the use away from the current document.

\section*{Usage}

\begin{texdisp}
\href{URL}{TEXT}
\end{texdisp}

This inserts the text \tex`TEXT` with a hyperlink to the web address \tex`URL`.

\section*{Example}

\begin{texdisp}
Scientific article preprints may be found on \href{https://arxiv.org}{the Ar$\chi$iv}.
\end{texdisp}

\section*{Remarks}

In Tectonopedia, this command simply inserts an \tex`<a>` HTML tag. It overrides
the command of the same name provided by hyperref.

The tag inserted by this command includes an attribute \tex`target="_blank"`,
causing the link to be opened in a new browser window or tab. The alternative
command \`@BhrefInternal` omits this.

\section*{See Also}

\begin{itemize}
\item \`@BhrefInternal`, which does not include the \tex`target="_blank"` attribute.
\end{itemize}


\Entry{hrefInternal}{\string\hrefInternal}{\hrefInternal}
\DeclareTerm*{\string\hrefInternal}{@BhrefInternal}

The Tectonopedia command \b{\string\hrefInternal} creates a hyperlink. It is
intended to be used for links internal to the current document.

\section*{Usage}

\begin{texdisp}
\hrefInternal{URL}{TEXT}
\end{texdisp}

This inserts the text \tex`TEXT` with a hyperlink to the web address \tex`URL`.

\section*{Example}

\begin{texdisp}
For more information, see \hrefInternal{../about/}{the About page}.
\end{texdisp}

\section*{Remarks}

The tag inserted by this command does not include any HTML attributes besides
\tex`href`. The alternative command \`@Bhref` is intended for outgoing
hyperlinks; it includes a \tex`target="_blank"` attribute as well, causing the
link to be opened in a new browser window or tab.

\section*{See Also}

\begin{itemize}
\item \`@Bhref`, intended for use with outgoing hyperlinks
\end{itemize}

